NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all

### 2023-01-26 main.tex
Update text

\begin{abstract}
We present a databank of lipid bilayer simulations from the NMRlipids open collaboration project.
\end{abstract}
\begin{document}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Figures/overlay2.pdf}
    \caption{a) Structure of an overlay databank. 
    %Raw data is available from a publicly accessible repository (layer 1).
    %The core of the databank contains information on the location of the raw data and indexing of molecular %compositions and simulation details described with universal naming conventions (layer 2).
    %The content of databank can be viewed and analysed using external programs that automatically read the information and data from layers 1 and 2 (layer 3). Examples of such applications are the GitHub repository generating the results presented in this work and the graphical access to the databank.
    A more detailed structure of the databank layer in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
    b) Distribution of the lengths of the trajectories, total number of trajectories and total length of the simulations in the NMRlipids databank.
    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or fewer simulations ('others') are listed on the right. 
    d) Currently available binary mixtures in the NMRlipids databank. 
    e) Distribution of force fields in the simulations in the NMRlipids databank.
    }
    \label{fig:overlay}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=88mm]{Figures/flowchart2.pdf}
    \caption{ A) Flowchart for analysing properties from all MD simulations in the NMRlipids databank.
    B) Flowchart for accessing the results analyzed from the NMRlipids databank.
    C) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields at 310\,K. The points from the best performing simulations in the NMRlipids databank quality evaluation for each lipid (Figs.~\ref{fig:quality} and~\ref{fig:POPC_POPE_dataSI}) are surrounded by black circles. 
    }
    \label{fig:POPCPOPEapls}
\end{figure}

\subsection{Selection of best simulation parameters for applications using the NMRlipids databank}
%Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations. 
Understanding of how closely simulations resemble the reality is crucial when assessing the reliability and significance of MD simulation results, as well as when using atomistic simulations in parametrization of coarse grained models. However, universal quality measures for MD simulations are lacking, and a prevailing force field that would perform best in all situations has not been found for membranes. For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}. On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}. Furthermore, Fig.~\ref{fig:POPCPOPEapls}B illustrates divergent predictions for the difference in packing between POPC and POPE membranes from different force fields. To streamline the selection of best simulations for specific applications, and to guide the development of atomistic and coarse grained parameters, quantitative quality measures for lipid bilayer simulations are defined in the NMRlipids databank. Together with the API, these will enable rapid and automatic quality evaluation of simulations in the NMRlipids databank, thereby promoting the selection of best simulations for specific applications without laborious force field evaluation, and automatic parametrization procedures for atomistic and coarse grained simulations.
%that can be used to rapidly find the best available parameters .

The quality for each simulation in the NMRlipids databank with the corresponding experimental data available is currently done using C--H bond order parameters from NMR experiments and form factors from X-ray scattering experiments. The C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, while correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F. The X-ray scattering form factors are related to membrane dimensions \textit{via} the electron density profile. In particular, locations of minima in these form factors are correlated with the membrane lateral packing and thickness as shown in Figs.~\ref{fig:quality}F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
For quantitative quality measures, probabilities for each order parameter to locate within experimental error, and the distance of the first form factor minimum from the experimental value are calculated from a simulation.
%For detailed definitions of quality measures, see the supplementary information.

Ranking based on C--H bond order parameter quality for the best performing simulations is shown in Fig.~\ref{fig:quality}A, and for the best performing POPE lipid bilayer simulations in Fig.~\ref{fig:quality}B. Direct comparisons with experiments are shown for selected simulations in Figs.~\ref{fig:quality}C--E. 
%The power of the NMRlipids databank in selecting the best simulation parameters for a specific application is demonstrated for mixtures of PE and PC lipids in Fig.~\ref{fig:POPCPOPEapls} showing area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K. Predictions from different force fields deviate significantly in terms of absolute values and the slope of decrease upon addition of POPE. 
This can be used to select the the simulations that best capture area per lipids for POPE and POPC lipids in Fig.~\ref{fig:POPCPOPEapls}B because the area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F).
%, simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation. 
Quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} reveal that Slipids ranks 1\textsuperscript{st} for POPC with a clear difference to others, and 2\textsuperscript{nd} for POPE with only marginally lower quality than GROMOS-CKP. Simulations with CHARMM36 and GROMOS-CKP predict overly packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while the area per lipid for POPC is overestimated in Lipid17. For POPE, CHARMM36 and Lipid17 predict membranes that are too packed. In conclusion, Slipids parameters is the best available choice for studies where packing effects of PE lipids are relevant

@@

\subsection{Detecting rare phenomena using NMRlipids databank: Cholesterol flip--flops}
Lipid flip--flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}. Phospholipid flip--flops are slow when not facilitated by proteins, occurring spontaneously on the timescale of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster. Still, the reported timescales range from minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}. These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, yet atomistic simulations reporting cholesterol flip--flop events have been published recently~\cite{gu19,javanainen19,baral20}. These studies report an increase in cholesterol flip--flop rates with increasing acyl chain unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but the amount of data in these individual studies was not sufficient to systematically assess correlations between cholesterol flip--flop rates and membrane properties. Here, we demonstrate that the analysis of such correlation is now accessible to all by using the API to access the large amount of MD simulation data in the NMRlipids databank as described in section~\ref{section:access}.

Using workflow depicted in Fig.~\ref{fig:POPCPOPEapls}A, we first calculated the cholesterol flip--flop rates from all the simulations available in the NMRlipids databank. Flip--flops were observed for cholesterol, % (634 events in 77 different simulations), 
DCHOL (18,19-di-nor-cholesterol), %, 16 flip-flops in 1 simulation) 
DOG (1,2-dioleoyl-\textit{sn}-glycerol), %, 44 flip-flops in 4 simulations) 
and SDG (1-stearoyl-2-docosahexaenoyl-\textit{sn}-glycerol) in the NMRlipids databank. Cholesterol flip--flop rates range between 0.001--1.6\,\textmu{}s$^{-1}$ with the mean of 0.16\,\textmu{}s$^{-1}$ and median of 0.07\,\textmu{}s$^{-1}$. These values are in line with the previously reported values from atomistic MD simulations~\cite{gu19,javanainen19,baral20}. The flip--flop rate of DCHOL, 0.2\,\textmu{}s$^{-1}$, was close to the average value of cholesterol, while average rates for diacylglycerols DOG and SDG were higher than for cholesterol, 0.4\,\textmu{}s$^{-1}$ and 0.5\,\textmu{}s$^{-1}$, respectively. Flip--flops were not observed for other lipids, giving the upper limits for PC lipid flip--flop rate of 9$\cdot$10$^{-6}$\,\textmu{}s$^{-1}$ and for ceramide (\textit{N}-palmitoyl-\textsc{D}-\textit{erythro}-sphingosine) of 0.002\,\textmu{}s$^{-1}$. Thus, available data in the NMRlipids databank suggests that the lipid flip--flip rate decreases in the order of diacylglycerols $>$ cholesterol $>$ other lipids including ceramides. However, the amount of data for diacylgycerols (8 simulations with Lipid17 force field) and ceramide (3 simulations with CHARMM36 force field) is less than for cholesterol (83 simulations), thus we cannot fully exclude the effect of force field or composition on this comparison.

Nevertheless, we use the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}B and the results for flip--flop rates from the NMRlipids databank to analyse how they depend on membrane properties. Cholesterol flip--flop rates and their histograms are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B--D. The results reveal a non-linear correlation between cholesterol flip--flop rate and membrane packing. Flip--flop rates increase by an order of magnitude when membrane packing density decreases and a major jump is observed at low membrane packing. The order of magnitude changes in cholesterol flip--flop rate with the membrane composition may have major implications in understanding lipid trafficking and membrane biochemistry~\cite{gu19,baral20}. Because the results from the NMRlipids databank are averaged over large range of membrane compositions and force fields, they show that the strong dependence of cholesterol flip--flop rate on membrane properties is not limited to certain lipid compositions or force fields used in previous studies~\cite{gu19,javanainen19,baral20}.

\begin{figure}[htb]
    \centering
    \includegraphics[width=88mm]{Figures/CholFlipFlops.pdf}
    \caption{\textbf{A} Illustration of cholesterol flip-flop.  
      \textbf{B--D} Cholesterol flip--flops analyzed from the databank as a function of membrane thickness, area per lipid, and acyl chain order. Values from simulations with non-zero flip--flop rates are shown with blue dots. Histogrammed values are shown with black dots.
    }
    \label{fig:flip-flops}
\end{figure}

Explorations of cholesterol flip--flops using MD simulations have only recently become possible for research groups with substantial resources and specialized expertise~\cite{gu19,javanainen19,baral20}. Our analysis demonstrates how the NMRlipids databank makes analyses of such rare events accessible to wide range of scientists in different fields of science and industry who do not have the access to required resources and expertise to produce the large amounts of MD simulation data required.

\subsection{Extending the scope of MD simulations to new fields using the NMRlipids databank: Water diffusion anisotropy in membrane systems}
The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes can be related to the translocation of drugs  through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and to the formation of signals in diffusion tensor MRI imaging~\cite{topgaard20}. MD simulations are rarely used to analyze the anisotropic diffusion of water since only few membrane permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of sufficient amount of data challenging. Here, we show that the API access to the data in NMRlipids databank enables systematic analyses on how the anisotropic diffusion of water depends on membrane properties in multilamellar membrane systems, thereby extending the applications of MD simulations to new fields.

@@

To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids databank using the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}A. The resulting non-zero values range between 0.3 and 322\,\textmu{}m/s with the mean and median of 14\,\textmu{}m/s and 8\,\textmu{}m/s, respectively. These values agree with the previously reported simulation results~\cite{venable19,camilo2022} and have the same order of magnitude as experimentally determined diffusive permeability coefficients, but are on average larger than the values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}. Using the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}b, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability} B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order. As expected, the permeability increases with the temperature, giving an average energy barrier of 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B. On the other hand, the permeability of water decreases on average when membranes become more packed, \textit{i.e.}, with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability} C--E). Permeation of water through bilayers also depends on membrane properties also according previous studies, but there is no established consensus on whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} is the main parameter determining the permeability. 
%is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}. 
Our analysis over the NMRlipids databank, containing significantly more data than what was available in previous studies, suggest non-linear dependencies on both of these parameters, yet the linear correlation would be a good approximation for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,\AA{}$^2$ (insets in Figs.~\ref{fig:permeability} C--D).
%within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
Clear dependencies of permeability on charged lipids, cholesterol, POPE, or hydration level were not observed in Fig.~\ref{fig:permeationSI} in the supplementary information.
%, although weak decrease for the latter two may be visible.
%
%
%
%%Several models on how water dynamics depends on membrane properties have been proposed based on different experimental and %%theoretical methods, yet consensus 
%%%on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, 
%%has not been reached~\cite{mathai07,nitsche13,nitsche16,shinoda16,venable19,frallicciardi22}.

@@

The focus of biomolecular simulations is moving from studies of individual molecules to larger complexes and even whole cells and organelles~\cite{johnson15,thornburg22,gupta22}. Simultaneously, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}. The NMRlipids databank will support the development in all these directions. The automatic quality evaluation and ranking of simulations in the NMRlipids databank guides the optimization of new force field parameters and the selection of best parameters for large biomolecular complexes where simulation quality becomes increasingly important due the accumulation of small errors in large simulations. Furthermore, the NMRlipids databank serves as a training set for diverse machine learning applications. For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments. In addition, the data contained in the NMRlipids databank will enable training of more elaborated models to predict arbitrary membrane properties.

Providing programmatic access to large scale MD simulation data in the NMRlipids databank can lead to applications in unprecedented directions in fields where MD simulations are less commonly utilized. Here, we exemplified this by analysing the anisotropic diffusion of water (Fig.~\ref{fig:permeability}), which is relevant in pharmacokinetic modeling and in MRI imaging, both areas where MD simulations are not yet commonly used~\cite{nitsche19,topgaard20}. 
%The NMRlipids databank will be of great benefit to scientists involved in MD simulations. They will be able to rapidly evaluate the quality of MD simulations in order to filter out potentially misleading results or facilitate force field parameter development~\cite{antila22b}. The selection of best models for a specific application is demonstrated for PC/PE lipid mixtures in Fig.~\ref{fig:quality}, and previously for PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}. 
The programmatic access in the NMRlipids databank enables automatic analyses over larger sets of simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) than is currently possible by a single research group. The added value that access to this simulation data provides is demonstrated here for the analysis of simulation qualities (Fig.~\ref{fig:quality}), water permeation through membranes (Fig.~\ref{fig:permeability}) and cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}). 
%Such models could be used to connect membrane composition to relevant membrane properties in wide range of fields from molecular biology to drug development and biomolecular imaging. 
Different types of applications enabled by the NMRlipids databank across a wide range of fields are exemplified in Table~\ref{tab:applications}. The ever growing amount of data is expected to further increase the scope of potential applications of the NMRlipids databank in biology, biotechnology,material science, biomolecular imaging and beyond.

>>>>

\begin{abstract}
Cellular membrane lipid composition is implicated in diseases and major biological functions in cells, but membranes are difficult to study experimentally due to their intrinsic disorder and complex phase behaviour. Molecular dynamics (MD) simulations have been useful in understanding membrane systems, but they require significant computational resources and often suffer inaccuracies in model parameters. Applications of data-driven and machine learning methods, currently revolutionising many fields, are limited to membrane systems due to the lack of suitable training sets. Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to quality evaluated atom-resolution MD simulations of lipid bilayers. The NMRlipids databank will benefit scientists in different disciplines by provoding automatic ranking of simulations based on their quality against experiments, flexible implementations of data-driven and machine learning applications, and rapid access to simulation data via graphical user interface. To demonstrate the unlocked possibilities beyond current MD simulation studies, we analyzed how anisotropic diffusion of water and cholesterol flip–flop rates depend on membrane properties.

\end{abstract}
\begin{document}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Figures/overlay2.pdf}
    \caption{a) Schematic presentation of the overlay structure used in the NMRlipids databank. A more detailed structure of the databank layer is shown in Fig.~\ref{DatabankStructure} in the SI.
    b) Distribution of the lengths of the trajectories, total number of trajectories and total length of the simulations in the NMRlipids databank.
    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or fewer simulations ('others') are listed on the right. 
    d) Currently available binary mixtures in the NMRlipids databank. 
    e) Distribution of force fields in the simulations in the NMRlipids databank.
    }
    \label{fig:overlay}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=88mm]{Figures/flowchart2.pdf}
    \caption{ A) Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids databank using the API.
    B) Flowchart for accessing results calculated from the NMRlipids databank and stored to the application layer.
    C) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields at 310\,K in simulations that are available in the NMRlipids databank. The points from the best performing simulations in the NMRlipids databank quality evaluation for each lipid (Figs.~\ref{fig:quality} and~\ref{fig:POPC_POPE_dataSI}) are surrounded by black circles. 
    }
    \label{fig:POPCPOPEapls}
\end{figure}

\subsection{Applications of the NMRlipids databank}
\subsubsection{Selection of simulation parameters using the NMRlipids databank: Best models for most abundant neutral membrane lipids}
%Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations. 
Understanding of how closely simulations resemble the reality is crucial for selecting best models for specific applications, assessing the reliability and significance of MD simulation results, and using atomistic simulations in parametrization of coarse grained models. However, undefined universal quality measures for MD simulations and the lack of a prevailing force field that would perform best in all simulations has led to many controversial results in membrane simulations~\cite{antila22b}. 
%For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}. On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}. 
For example, Fig.~\ref{fig:POPCPOPEapls}B illustrates divergent predictions from different force fields for the lateral packing of POPC and POPE membranes that are the two most biologically abundant neutral membrane lipids~\cite{vanmeer08}. We have previously resolved such controversies for membrane structures and ion binding affinities for neutral~\cite{botan15,catte16} and charged lipids~\cite{antila19,bacle21} by comparing simulations with C-H bond order parameters from NMR, yet this requires manual comparisons between large number of simulations. To streamline this process,
%the selection of best simulations for specific applications, and to guide the development of atomistic and coarse grained parameters,
we have defined quantitative quality measures, based on C--H bond order parameters from NMR and X-ray scattering form factors, for lipid bilayer simulations in the NMRlipids databank. 
%The quality for each simulation in the NMRlipids databank with the corresponding experimental data available is currently done using C--H bond order parameters from NMR experiments and form factors from X-ray scattering experiments. 
The C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, while correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F. The X-ray scattering form factors are related to membrane dimensions \textit{via} the electron density profile. In particular, locations of minima in these form factors are correlated with the membrane lateral packing and thickness as shown in Figs.~\ref{fig:quality}F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}. For quantitative quality measures, probabilities for each order parameter to locate within experimental error, and the distance of the first form factor minimum from the experimental value are calculated from a simulation.
%For detailed definitions of quality measures, see the supplementary information.
Together with the API, these enable rapid and automatic quality evaluation of simulations in the NMRlipids databank, thereby promoting the selection of best simulations for specific applications without laborious force field evaluation, and automatic parametrization procedures for atomistic and coarse grained simulations.

To select the best models for POPC and POPE among simulations compared in Fig.~\ref{fig:POPCPOPEapls}C, we show the ranking based on C--H bond order parameter quality in Fig.~\ref{fig:quality}A for the best performing simulations in the NMRlipids databank, and for the best performing POPE lipid bilayer simulations in Fig.~\ref{fig:quality}B. Direct comparisons with experiments are shown for selected simulations in Figs.~\ref{fig:quality}C--E and~\ref{fig:POPC_POPE_dataSI}. 
%The power of the NMRlipids databank in selecting the best simulation parameters for a specific application is demonstrated for mixtures of PE and PC lipids in Fig.~\ref{fig:POPCPOPEapls} showing area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K. Predictions from different force fields deviate significantly in terms of absolute values and the slope of decrease upon addition of POPE. 
Because the area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} can be used to select the simulations that best capture area per lipids for POPE and POPC lipids in Fig.~\ref{fig:POPCPOPEapls}C. In these results, simulations with Slipids parameters ranks 1\textsuperscript{st} for POPC with a clear difference to others, and 2\textsuperscript{nd} for POPE with only marginally lower quality than GROMOS-CKP. Simulations with CHARMM36 and GROMOS-CKP parameters predict overly packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while the area per lipid for POPC is overestimated in Lipid17. For POPE, CHARMM36 and Lipid17 predict membranes that are too packed. In conclusion, the quality evalution based on the NMRlipids databank suggests that Slipids parameters are the best available choice for simulations with PC and PE lipids, at least for the cases were membrane packing is relevant.

@@

\subsubsection{Detecting rare phenomena using NMRlipids databank: Cholesterol flip--flops}
Lipid flip--flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}. Phospholipid flip--flops are slow when not facilitated by proteins, occurring spontaneously on the timescale of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster. Still, the reported timescales range from minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}. These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, yet atomistic simulations reporting cholesterol flip--flop events have been published only recently~\cite{gu19,javanainen19,baral20}. These studies report an increase in cholesterol flip--flop rates with increasing acyl chain unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but the amount of data in these individual studies was not sufficient to systematically assess correlations between cholesterol flip--flop rates and membrane properties. Here, we demonstrate that the NMRlipids databank API makes analyses of such rare phenomena accessible to everyone by enabling the access to large amount of MD simulation data as described in section~\ref{section:access}. This is particularly useful for scientists in different fields of science and industry who do not have the access to required computational resources and expertise to produce the large amounts of MD simulation data.

Using workflow depicted in Fig.~\ref{fig:POPCPOPEapls}A, we first calculated the cholesterol flip--flop rates from all the simulations available in the NMRlipids databank. Flip--flops were observed for cholesterol, % (634 events in 77 different simulations), 
DCHOL (18,19-di-nor-cholesterol), %, 16 flip-flops in 1 simulation) 
DOG (1,2-dioleoyl-\textit{sn}-glycerol), %, 44 flip-flops in 4 simulations) 
and SDG (1-stearoyl-2-docosahexaenoyl-\textit{sn}-glycerol). The observed cholesterol flip--flop rates, ranging between 0.001--1.6\,\textmu{}s$^{-1}$ with the mean of 0.16\,\textmu{}s$^{-1}$ and median of 0.07\,\textmu{}s$^{-1}$, are in line with the previously reported values from atomistic MD simulations~\cite{gu19,javanainen19,baral20}. The flip--flop rate of DCHOL, 0.2\,\textmu{}s$^{-1}$, was close to the average value of cholesterol, while average rates for diacylglycerols DOG and SDG were higher than for cholesterol, 0.4\,\textmu{}s$^{-1}$ and 0.5\,\textmu{}s$^{-1}$, respectively. Flip--flops were not observed for other lipids, giving the upper limits for PC lipid flip--flop rate of 9$\cdot$10$^{-6}$\,\textmu{}s$^{-1}$ and for ceramide (\textit{N}-palmitoyl-\textsc{D}-\textit{erythro}-sphingosine) of 0.002\,\textmu{}s$^{-1}$. Thus, available data in the NMRlipids databank suggests that the lipid flip--flip rate decreases in the order of diacylglycerols $>$ cholesterol $>$ other lipids including ceramides. However, the amount of data for diacylgycerols (8 simulations with Lipid17 force field) and ceramide (3 simulations with CHARMM36 force field) is less than for cholesterol (83 simulations), thus we cannot fully exclude the effect of force field or composition on this comparison.

Nevertheless, we used the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}B to analyse how flip--flop rates calculated from the NMRlipids databank depend on membrane properties. Cholesterol flip--flop rates and their histograms are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B--D. The results reveal a non-linear correlation between cholesterol flip--flop rate and membrane packing which is depicted as area per lipid. Flip--flop rates increase by an order of magnitude when membrane packing density decreases and a major jump is observed at low membrane packing. The order of magnitude changes in cholesterol flip--flop rate with the membrane composition may have major implications in understanding lipid trafficking and membrane biochemistry~\cite{gu19,baral20}. Because the results from the NMRlipids databank are averaged over large range of membrane compositions and force fields, they show that the strong dependence of cholesterol flip--flop rate on membrane properties is not limited to certain lipid compositions or force fields used in previous studies~\cite{gu19,javanainen19,baral20}.

\begin{figure}[htb]
    \centering
    \includegraphics[width=88mm]{Figures/CholFlipFlops.pdf}
    \caption{\textbf{A} Illustration of cholesterol flip--flop.  
      \textbf{B--D} Cholesterol flip--flops analyzed from the databank as a function of membrane thickness, area per lipid, and acyl chain order. Values from simulations with non-zero flip--flop rates are shown with blue dots. Histogrammed values are shown with black dots.
    }
    \label{fig:flip-flops}
\end{figure}

\subsubsection{Extending the scope of MD simulations to new fields using the NMRlipids databank: Water diffusion anisotropy in membrane systems}
The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes can be related to the translocation of drugs  through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and to the formation of signals in diffusion tensor MRI imaging~\cite{topgaard20}. MD simulations are rarely used to analyze the anisotropic diffusion of water since only few membrane permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of sufficient amount of data challenging. Here, we show that the API access to the data in NMRlipids databank enables systematic analyses on how the anisotropic diffusion of water depends on membrane properties in multilamellar membrane systems, thereby extending the applications of MD simulations to new fields.

@@

To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids databank using the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}A. The resulting non-zero values range between 0.3 and 322\,\textmu{}m/s with the mean and median of 14\,\textmu{}m/s and 8\,\textmu{}m/s, respectively. These values agree with the previously reported simulation results~\cite{venable19,camilo2022}, but are on average slightly larger than experimental values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}. Using the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}B, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability}B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order. As expected, the permeability increases with the temperature, giving an average energy barrier of 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability}B. On the other hand, the permeability of water decreases on average when membranes become more packed, \textit{i.e.}, with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability} C--E). Permeation of water through bilayers depends on membrane properties also according to previous studies, but there is no established consensus on whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} is the main parameter determining the permeability. 
%is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}. 
Our analysis over the NMRlipids databank, containing significantly more data than what was available in previous studies, suggest non-linear dependencies on both of these parameters, yet the linear correlation would be a good approximation for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,\AA{}$^2$ (insets in Figs.~\ref{fig:permeability} C--D).
%within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
Clear dependencies of permeability on charged lipids, cholesterol, POPE, or hydration level were not observed in Fig.~\ref{fig:permeationSI} in the supplementary information.
%, although weak decrease for the latter two may be visible.
%
%
%
%%Several models on how water dynamics depends on membrane properties have been proposed based on different experimental and %%theoretical methods, yet consensus 
%%%on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, 
%%has not been reached~\cite{mathai07,nitsche13,nitsche16,shinoda16,venable19,frallicciardi22}.

@@

The focus of biomolecular simulations is moving from studies of individual molecules to larger complexes and even whole cells and organelles~\cite{johnson15,thornburg22,gupta22}. Simultaneously, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}. The NMRlipids databank will support the development in both of these directions. The automatic quality evaluation and ranking of simulations in the NMRlipids databank guides the optimization of new force field parameters and the selection of best parameters for large biomolecular complexes where simulation quality becomes increasingly important due the accumulation of small errors in large simulations. Furthermore, the NMRlipids databank serves as a training set for diverse machine learning applications. For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments. In addition, the data contained in the NMRlipids databank will enable training of more elaborated models to predict arbitrary membrane properties.

%The NMRlipids databank will be of great benefit to scientists involved in MD simulations. They will be able to rapidly evaluate the quality of MD simulations in order to filter out potentially misleading results or facilitate force field parameter development~\cite{antila22b}. The selection of best models for a specific application is demonstrated for PC/PE lipid mixtures in Fig.~\ref{fig:quality}, and previously for PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}. 
The programmatic access in the NMRlipids databank enables automatic analyses of structural and dynamic properties over larger sets of simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) than is currently possible by a single research group. The added value of such analyses is demonstrated here by analysing the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}) and cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}). By enabling these analyses for everyone, the NMRlipids databank can lead to applications in unprecedented directions in fields where MD simulations are less commonly utilized, such as pharmacokinetic modeling and in MRI imaging where anisotropic diffusion of water play important roles~\cite{nitsche19,topgaard20}. 
Different types of applications enabled by the NMRlipids databank across a wide range of fields are exemplified in Table~\ref{tab:applications}. The ever growing amount of data is expected to further increase the scope of potential applications of the NMRlipids databank in biology, biotechnology, material science, biomolecular imaging and beyond.

\begin{table}[t]
    \centering
    \begin{tabular}{p{5.0cm}  p{5.0cm}  p{4.0cm}}
    Type of application     & Practical examples & Target group \\
    \hline
    Analyses of rare phenomena               & Lipid flip--flops, water permeation (Figs.~\ref{fig:flip-flops} and~\ref{fig:permeability}) & Membrane scientists \\
    \\
    Correlations between membrane properties & 
    Membrane structural properties, water dynamics (Figs.~\ref{fig:quality} and~\ref{fig:permeability}) & 
    Membrane scientists \\
    \\
    Applications that are outside typical scope of MD simulations & 
    Anisotropic water diffusion for pharmacokinetics and MRI imaging applications (Fig.~\ref{fig:permeability}) & 
    Scientist in fields where MD simulations are not usually applied \\
    \\
    Selection of the best simulation model for a specific application & 
    Best model for POPC lipids (Fig.~\ref{fig:quality}), headgroup conformations~\cite{bacle21}, 
    packing of PS~\cite{antila22b} and PE (Fig.~\ref{fig:POPCPOPEapls}) containing membranes. &
    Scientists using MD simulations \\
    \\
    Guidance for force field development & 
    Improvements in ion binding to lipids~\cite{melcr18,melcr20} and lipid headgroup conformational ensembles~\cite{yu21,dickson22,grote20} &
    Scientists developing parameters for MD simulations \\
    \\
    Training and target data for coarse grained models & 
    Optimizing parameters of coarse grained models against NMRlipids databank, extracting continuum parameters for membranes. &
    Scientists developing and using coarse grained MD simulations \\
    \\
    Training set for machine learning applicatons &
    Programmatic access to the data and results enables training of machine learning type of models for various applications, such as predictions of membrane properties from composition & Scientists building and using machine learning applications for biomolecules.  
    \end{tabular}
    \caption{Examples on applications of the NMRlipids databank.}
    \label{tab:applications}
\end{table}
